---
title: "Prediction in R"
author: "Tom Mathews"
date: "October 29, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## R Markdown
The goal is to create a model that can predict if users are performing their excercises correctly.

My initial thought was that I would leverage the temporal nature of the sensors to identify cycles of movement. 
However, upon review of the test set (in which the data isn't necessarily sequential in time), it became clear 
that the desired models would need to support sampling the sensors (rather than modeling continuous movement).

## Cleaning Data
The data had a lot of gaps in it that requried cleaning:
```{r cars}
library(caret)
library(lubridate)
library(dplyr)

setwd("C:\\Users\\Tom\\Downloads\\Coursera\\Practical Machine Learning\\")

#read the files
Q4.training <- read.csv("pml-training.csv", na.strings = c('', 'NA', '#DIV/0!'))
Q4.testing <- read.csv("pml-testing.csv", na.strings = c('', 'NA', '#DIV/0!'))

```

Then, I wanted to remove the NA columns, and focus on columns with higher variance:
```{r}
Q4.training_NoNACols <- Q4.training[, colSums(is.na(Q4.training)) != nrow(Q4.training)]
Q4.NewTraining <- select(Q4.training_NoNACols, matches("classe|.+_(roll|pitch|yaw)_.+"), -amplitude_yaw_forearm, -amplitude_yaw_dumbbell, -amplitude_yaw_belt, -nearZeroVar(Q4.training_NoNACols)) 
inTrain = createDataPartition(y = Q4.NewTraining$classe, p = 0.8, list = FALSE)
Q4.NewTesting <-  Q4.NewTraining[-inTrain, ]

```


## Creating the model

I created a random forest model:

```{r pressure, echo=FALSE}
set.seed(42)
Q4.rf <- train(classe ~ . , method="rf", data=Q4.NewTraining, na.action=na.omit)

print(Q4.rf)
```

Unfortunately, at this time I ran into two problems in R that I couldn't find a solution for. :(

First, I was unable to use the testing data to validate the model (hence splitting it manually from the training data up above). 

Secondly, the confusion matrix can't be generated, as the number of predictions is always much smaller than the number of values passed into the prediction algorithm. This causes the confusion matrix to fail.

```{r arg}
Q4.rf_Predict <- predict(Q4.rf, newdata=Q4.NewTesting)

#Q4.rf_Confusion <- confusionMatrix(Q4.NewTesting$classe, Q4.rf_Predict)

#print(Q4.rf_Confusion)
```